---
title: "Assignment Part III"
author: "Silpa Soni Nallacheruvu (19980824-5287) Hernan Aldana (20000526-4999)"
date: "`r Sys.Date()`"
output: pdf_document
---

set.seed(980824)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE}
source("../Codes/Functions.R")
source("codes_assignment_3.R")
```

# Summary

This report covers three main tasks under a Bayesian framework applied to 
logistic regression modeling. Task 1 focuses on validating AIC calculations and leave-one-out cross-validation to assess model accuracy. Task 2 involves 
implementing a function to compute the posterior density of regression 
parameters, enhancing the model’s estimation under Bayesian principles.
Task 3 is dedicated to applying the Metropolis-Hastings algorithm for 
posterior sampling, aiming to simulate from the posterior distribution of 
parameters, assess convergence, and estimate credibility intervals 
and predictive probabilities for new observations.

# Task 1

## Approach :

To calculate the AIC, the formula

$2*k - 2*l(\hat{\theta_{ml}})$

was applied where k is the number of parameters in the theta. To calculate the leave-one-out cross-validation, the formula

$\frac{\sum_{i=1}^{n} {l_i(\hat{theta_i)}}}{n}$

was applied where $\hat{\theta_i}$ is $\hat{\theta_{ml}}(X_{-i})$, $(X_{-i})$ is one observation $(X_{i})$ left out of X, $l_i(\hat{theta_i)}$ is the log likelihood for the i-th left out observation and n is the total number of observations. These values were then compared with the AIC from R in the provided summary.

## Code :

```{r echo=TRUE}
# ---- Task_1 ----

# Compute AIC = 2k - 2l(theta_ml)
n <- nrow(X)
theta0 = rep(0, ncol(X))
k <- length(theta0)
theta_estimate <- NR(theta0, 3, y, X) 
log_likelihood <- l(theta_estimate, y, X)
aic_computed <- 2*k - 2*log_likelihood

#AIC output from R summary 
r_summary_aic <- summary(modell)$aic

# Compute k_cv =sum(l_i(theta_i))/n
# Here, theta_i = theta_ml(X_-i)
nk_cv <- 0
for(i in 1:n) {
  # remove the i-th row from X and return a matrix
  X_minus_i <- X[-i, , drop=FALSE]
  y_minus_i <- y[-i, , drop=FALSE]
  X_i <- X[i, , drop=FALSE]
  y_i <- y[i, , drop=FALSE]
  theta_i <- NR(theta0, 3, y_minus_i, X_minus_i)
  # log likelihood for i-th observation
  log_likelihood_theta_i <- l(theta_i, y_i, X_i)
  nk_cv <- nk_cv + log_likelihood_theta_i
}

k_cv <- nk_cv/n

# Creating a comparison data frame 
comparison_aic_values <- data.frame(
  "AIC_R_model" = r_summary_aic,
  "AIC_computed" = aic_computed,
  "2*nK_CV_computed" = 2*nk_cv
)
```

## Output :

```{r}
comparison_aic_values
```

## Observation :

The computed AIC value and the AIC value from the R summary coincide in value. The computed K_CV is in the same magnitude as $AIC/(-2n)$ as expected.

# Task 2

## Approach :

The a posteriori density combines the a priori and the likelihood of the data. For the logistic regression, we have y data, X and the parameter vector $\theta$. so the a posteriori density is proportional to: $P(\theta|y,X) \propto P(y|X,\theta)P(\theta)$ where $P(\theta)$ is the a prior density and $P(y|X,\theta)$ is. the likelihood from the logistic regression model.

for a binary outcome, the likelihood for the logistic regression is given by the following equation:

$P(y_{i}|X_{i},\theta) =  \frac{1}{(1+\exp(-X_{i}\theta)}$

The a priori is Gaussian: $\theta \sim N(0, 100I)$ so the a prior density is the following: $P(\theta) \propto exp(-\frac{1}{2}\theta^T(100I)^{-1}\theta)$

## Code :

```{r echo=TRUE}
# ---- Task_2 ----
post <- function(theta, y, X) {
  eta <- X %*% theta # Logistic regression likelihood
  likelihood <- prod(plogis(eta)^y * (1 - plogis(eta))^(1 - y))
  prior <- exp(-0.5 * sum(theta^2 / 100)) # a priori with theta ~ N(0, 100 * I)
  posterior <- likelihood * prior # posteriori is proportional to a priori * likelihood
  return(posterior)
}

```

## Output :

```{r}
Xtest <- cbind(1, 18:25, rep(c(0, 1), 4), rep(c(1, 1, 0, 0), 2))
ytest <- c(rep(TRUE, 4), rep(FALSE, 4))
testing<-post(c(260, -10, 10, -20), ytest, Xtest) / post(c(270, -15, 15, -25), ytest , Xtest)
testing
```

## Observation :

Given the results obtained by the test the function works correctly.

# Task 3

## Approach :

Implemented the Metropolis-Hastings algorithm over 10000 iterations while
checking on the condition if the acceptance probability is greater than 
the acceptance ratio which is generated randomly as $ratio \sim U(0,1)$. 
We accepted the theta proposal if the condition was true. 
The acceptance probability is the ratio of the densities of the proposal
distribution of $\theta^\star$ and $\theta^{t-1}$.
The proposal distribution here is the previously computed 
posterior distribution.

## Code :

```{r echo=TRUE}

# ---- Task_3 ----

N <- 20000 
mh_algo <- function(theta_estimate, y, X) {
  theta <- matrix(nrow = N, ncol = 4)
  # initial value as the calculated theta_estimate
  theta[1,] <- theta_estimate 
  # here, suggested sigma as standard error from part II
  sigma <- standard_error(theta_estimate, y, X) 
  for (i in 2:N) {
    theta_star <- theta[i-1,] + rnorm(4) * sigma
    # check if the acceptance probability is greater than the acceptance ratio
    posterior_theta_star <- post(theta_star, y, X) 
    posterior_theta <- post(theta[i-1,], y, X)
    u <- runif(1)
    acceptance_probability <- min(1, posterior_theta_star/ posterior_theta)
    if (posterior_theta_star/ posterior_theta >= u) {
      theta[i,] <- theta_star  # Accept the proposal
    } else {
      theta[i,] <- theta[i-1,]  # Reject the proposal and fall back
    }
  }
  return(theta)
}

theta_sample <- mh_algo(theta_estimate, y, X)
```

## Output :

### Parameter Plots

```{r fig.show='hold', fig.align='center', fig.width=12, fig.height=6}
par(mfrow=c(2,2))
plot(x = c(1:N), y = theta_sample[,1], type = "l", col = "black",
main="theta1 Parameter Plot", xlab= "#iterations", ylab = "theta1") 

plot(x = c(1:N), y = theta_sample[,2], type = "l", col = "black", 
main="theta2 Parameter Plot", xlab= "#iterations", ylab = "theta2") 

plot(x = c(1:N), y = theta_sample[,3], type = "l", col = "black", 
main="theta3 Parameter Plot", xlab= "#iterations", ylab = "theta3") 

plot(x = c(1:N), y = theta_sample[,4], type = "l", col = "black", 
main="theta4 Parameter Plot", xlab= "#iterations", ylab = "theta4") 
```

### Parameter Posteriors Histograms

```{r fig.show='hold', fig.align='center', fig.width=5, fig.height=7}
par(mfrow=c(2,2))
hist(x = theta_sample[(1:N), 1], col = "lightgray",
main="theta1 Posterior Histogram", xlab= "theta1", ylab = "frequency") 

hist(x = theta_sample[(1:N), 2],  col = "lightgray", 
main="theta2 Posterior Histogram", xlab= "theta2", ylab = "frequency") 

hist(x = theta_sample[(1:N), 3], col = "lightgray", 
main="theta3 Posterior Histogram", xlab= "theta3", ylab = "frequency") 

hist(x = theta_sample[(1:N), 4], col = "lightgray", 
main="theta4 Posterior Histogram", xlab= "theta4", ylab = "frequency") 
```

### 95% credibility intervals

```{r}
c_i <- apply(theta_sample, 2, quantile, probs = c(0.025, 0.975))
print(t(c_i))

# frequentist 95% ci:
frequentist_ci <- cbind(theta_estimate- 1.96 * standard_error(theta_estimate, y, X),
                        theta_estimate + 1.96 * standard_error(theta_estimate, y, X))
print(frequentist_ci)
```

### Future Test

We calculate the conditional probability of success as $(P(Y^* = 1 \mid y) = E(p(x^*) \mid y))$, where $(x^*)$ represents the features of the specific observation as 
mentioned. 

Here, $p(x^*)$ is a function of random variable $\theta$. 
Used the logistic formula : $p(x^*, \theta) = \frac{1}{1 + \exp(-\theta^T x^*)}$

Hence, $E(p(x^*, \theta) \mid y))$ is calculated using Monte Carlo Integration 
by the sample mean $\frac{1}{N} \sum_{i=1}^N p(\theta_i, x^*)$ 
for N = 10000 samples and $\theta_i$ as i-th sample from theta estimate 
of MH algorithm.



```{r}
# a privately educated subject of your own sex and age.
# intercept = 1 
# Alder = 26    # Own age
# KonMan = 0    # Female
# Utbildare = 0 # Privately educated
X_star <- matrix(c(1, 26, 0, 0), nrow = 4, ncol = 1)
# model the probability using regression framework
p_x <- function(theta, x) {
  return (1 / (1 + exp(theta %*% x)))
}
# compute the probability of pass wrt X_star
prob_y <- rep(0, N)
for(i in 1:N) {
  theta <- theta_sample[i, , drop=FALSE]
  prob_y[i] <- p_x(theta, X_star)
}
expected_prob_y <- mean(prob_y)
cat("expected probability of y* : ", expected_prob_y)
```
